Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-30 evaluation.tex
Updated area and power results.

We implemented our monitoring architecture for adjustable overheads by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We model the main and monitoring cores as running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for flags.

@@

% Area and Power Overheads
\begin{table}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/area_power}
    \end{footnotesize}
    \caption{Average power overhead for dropping hardware at a 10\% overhead
    budget. Percentages in parentheses are normalized to the main core
    power.}
    \vspace{-0.2in}
    \label{tab:evaluation.area_power}
  \end{center}
\end{table}

Adding the dataflow engine in order to enable filtering and adjustable adjustable adds
overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
a first-order estimate of these area and power overheads in a 40 nm technology
node. McPat estimates the main core area as 1.96 mm$^2$ and the peak power usage as
121 mW averaged across all benchmarks. The average runtime power usage was 38.2 
mW. These area and power numbers consist of the core and
L1 cache, but do not include L2 cache, memory controller, and other
peripherals. The power numbers include dynamic as well as static (leakage)
power. For the dataflow engine, we modeled the ALUs used for address
calculation, the dataflow flag register file and cache, the configuration
tables, and the filter decision table. These were modeled using the
corresponding objects in McPat. We
note that this is only a rough area and power result since components such as the
wires connecting these modules have not been modeled. However, this gives a
sense of the order-of-magnitude overheads involved with implementing our
approach.

Our results show that an additional 0.183 mm$^2$ of silicon area is needed, an
increase of 9\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads. The peak power is 13-19 mW, which is
11-16\% of the main core's peak power usage. The average runtime power is 8-11
mW, corresponding to 21-29\% of the main core's runtime power.

>>>>

We implemented dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We model the main and monitoring cores as running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for flags.

@@

% Area and Power Overheads
\begin{table}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/area_power}
    \end{footnotesize}
    \caption{Average power overhead for dropping hardware at a 50\% overhead
    budget. Percentages in parentheses are normalized to the main core
    power.}
    \vspace{-0.2in}
    \label{tab:evaluation.area_power}
  \end{center}
\end{table}

Adding the dataflow engine in order to enable filtering and adjustable adjustable adds
overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
a first-order estimate of these area and power overheads in a 40 nm technology
node. McPat estimates the main core area as 2.71 mm$^2$ and the peak power usage as
965 mW averaged across all benchmarks. The average runtime power usage was 385 
mW. These area and power numbers consist of the core and
L1 cache, but do not include L2 cache, memory controller, and other
peripherals. The power numbers include dynamic as well as static (leakage)
power. For the dataflow engine, we modeled the ALUs used for address
calculation, the dataflow flag register file and cache, the configuration
tables, and the filter decision table. These were modeled using the
corresponding memory and ALU objects in McPat. We
note that this is only a rough area and power result since components such as the
wires connecting these modules have not been modeled. However, this gives a
sense of the order-of-magnitude overheads involved with implementing our
approach.

Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
shows the peak and runtime power overheads averaged across all benchmarks with
a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
5-7\% of the main core's peak power usage. The average runtime power is 3-4 
mW, corresponding to 8-11\% of the main core's runtime power.
